---
title: "Assignment 3"
author: "NGUYEN Ngoc Nhu Y"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "",
                      message = FALSE,
                      warning = FALSE)
```

# Task 1
```{r}
#Read the dataset and replace "?" with NA
engine <- read.csv("Engine.csv", na.strings = "?")
automobile <- read.csv("Automobile.csv", na.strings = "?")
maintenance <- read.csv("Maintenance.csv", na.strings = "?")
```

```{r}
#Inspect the structure of the dataset "Engine.csv"
str(engine)
#Inspect the structure of the dataset "Automobile.csv"
str(automobile)
#Inspect the structure of the dataset "Maintenance.csv"
str(maintenance)
```

```{r}
#Count total missing values after replacing "?" with NA
cat("Missing values in Engine:", sum(is.na(engine)), "\n")
cat("Missing values in Automobile:", sum(is.na(automobile)), "\n")
cat("Missing values in Maintenance:", sum(is.na(maintenance)), "\n")
```

```{r}
# Check which columns contain NA in the Engine dataset
colSums(is.na(engine))
# Count number of rows in Engine affected by missing values
rows_with_na <- sum(rowSums(is.na(engine)) > 0)
cat("Rows with missing values in Engine:", rows_with_na, "\n")
```
## **Notes:**

Only Engine has missing values "?" (Engine = 6, Automobile & Maintenance
= 0).

-\> only Engine has affected rows (Missing values occur only in
Horsepower)

-\> only Engine needs cleaning

When inspecting the column, it is appeared that there are:

-   5 missing entries occur in EngineType

-   1 missing entry occurs in Horsepower

**Replacing "?" with NA does not change the distribution because no
actual values are modified; NA only marks missing entries.**

```{r}
#Convert categorical variables to factors
automobile$BodyStyles <- as.factor(automobile$BodyStyles)
engine$FuelTypes <- as.factor(engine$FuelTypes)
maintenance$ErrorCodes <- as.factor(maintenance$ErrorCodes)

#Compute the median Horsepower for imputation
median_hp <- median(engine$Horsepower, na.rm = TRUE)
```

